---
title: "Problem Set 9"
author: "Lionel Chambon"
date: "2024-04-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

library(haven)
library(lmtest)
library(dplyr)
library(AER)
```

# Empirical

## Question 6

```{r cars}
fulton <- read_dta("fulton.dta")

# Refer to the theory pdf for the derivations of the reduced form equations.

qty_reduced2 <- lm(qty ~ stormy, data = fulton)

summary(qty_reduced2)


# Proceeding similarly for price:

price_reduced2 <- lm(price ~  stormy, data = fulton)

summary(price_reduced2)

```

```{r}

coef1 <- summary(qty_reduced2)$coefficients[2]
coef2 <- summary(price_reduced2)$coefficients[2]

coef_interpretation1 <- exp(summary(qty_reduced2)$coefficients[2]) - 1
print(coef_interpretation1)

coef_interpretation2 <- exp(summary(price_reduced2)$coefficients[2]) - 1
print(coef_interpretation2)

```


The equation for quantity and the coefficient yielded tells us stormy is 
negatively correlated with supply. Since prices and quantities are expressed
in log according to an online description of the dataset, exponentiating and 
subtracting 1 yields an approx. 30.4% decrease in supply relative to when
there is no storm.

Similarly, is is positively correlated with the price. Exponentiating and 
subtracting 1 yields an approx. 39,8% increase in price relative to when there
is no storm.

## Question 7

```{r}

ratio1 <- coef1/coef2
print(ratio1)

```

We are estimating $\alpha_2$ (isolating $z$ in the reduced form equations and then dividing the coefficients will leave $\alpha_2$). We are thus looking at the effect of prices on demand. Since prices and quantities are expressed in logs, this shows us that a 1% price increase leads to a demand decrease of about 1,08%.

## Question 8

```{r}

# Since we are told to use the equivalent of ivreg, I decided to use the R
# equivalent as opposed to both 2SLS regressions like last week.

ivmodel <- ivreg(qty ~ price | stormy, data = fulton)

summary(ivmodel)

ivmodel2 <- ivreg(price ~ qty | stormy, data = fulton)

summary(ivmodel2)

coef3 <- exp(summary(ivmodel)$coefficients[2]) - 1
print(coef3)

coef4 <- exp(summary(ivmodel2)$coefficients[2]) - 1
print(coef4)

ratio2 <- coef3/coef4
print(ratio2)

```

The coeffient is identical to the previous result.

## Question 9

```{r}

# Now repeating the exercise with the coefficient for friday:

qty_reduced_f <- lm(qty ~ + stormy + friday, data = fulton)
summary(qty_reduced_f)


# Proceeding similarly for price:

price_reduced_f <- (lm(price ~ stormy + friday, data = fulton))
summary(price_reduced_f)

```

According to this output, the effect of friday is not significant. This could be a reason why we excluded it. Since we want to estimate the coefficient for price, friday may not be a good instrument since it does not produce any meaningul exogenous variation in price that we could use as instruments, unlike a storm.






